### 2010-09-05 src/Documentation/Introduction.tex
a start at adding some opeing waffle to the Epitome Intro

\section{In the beginning}

We really should write something informative here.

>>>>

\emph{''I believe that the time is ripe for significantly better
  documentation of programs, and that we can best achieve this by
  considering programs to be works of literature.''} Donald Knuth,
1992.

This document is intended to realise this lofty aim for the
implementation of Epigram. Although, given that it is being written by
several different people at once, the plot changes constantly, and
it's currently pushing 300 pages even though it has barely been started,
I don't expect it'll be a best seller.

Epigram is a many faceted beast whose facets are constructed roughly
one on top of the other. This document starts by describing its
bowels: the core type theory which serves as Epigram's language of
evidence and into which programs are elaborated. We then proceed
outwards and upwards until we reach its snorting, wild eyed, and rabid
face, a.k.a. the high-level language.

\section{In the beginning}
In the beginning there was Epigram 1. The Epigram 1 language was
designed by Conor McBride and James McKinna in their weighty paper
``The view from the left''\cite{conor.james:viewfromleft}. It descibes
a programming language with support for inductive families and
dependent pattern matching. Notably pattern matching is an added extra
to the type theory.  Instead, pattern matching programs in the high
level language are \emph{elaborated} into type theoretic expressions
made of traditional eliminators in the underlying theory. An
implementation of Epigram 1 was written by Conor and with its
idiosyncratic emacs interface with 2d syntax, one could make small
experiments. Unfortunately this prototype implementation proved
unscalable and unmaintainable. Since then Agda 2 and Coq's Russell
language have assimilated many of Epigram 1's features and many of
Epigram's programs can be written in these two other languages.

Epigram 2 is not just an attempt at a better implementation of Epigram
1; it has a radically different underlying type theory which supports
interesting features in the high-level language. Firstly, it supports
extensional equality of functions in an intentional theory by using
observational equality. Secondly, it is a closed type theory where new
data definitions are new codes in a universe of datatype descriptions
whose validity is internally guaranteed rather than decreed to be
acceptable by an external checker.
